---
title: "Assignment 4 - RMF Analysis"
author: "Philipp Markopulos"
date: "November 28th, 2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{titling}
- \posttitle{\end{center}}
---

```{r setup}
load(url("http://statmath.wu.ac.at/~vana/datasets/RMFAnalysisClean.rda"))
head(RMFAnalysisClean)
```

Recency (How recently did the customer make a purchase?); Frequency (How often do they purchase); Monetary Value (How much do they spend?)

# Visualisation
## 2D graphs

```{r cars}
#plotly used for visualisation

# make 2d pairs
op <- par(mfrow = c(1,3))
plot(RMFAnalysisClean$Recency, RMFAnalysisClean$Frequency, 
     ylab = "Frequency", xlab = "Recency")
plot(RMFAnalysisClean$Recency, RMFAnalysisClean$Monetary.mean, 
     ylab = "monetary mean", xlab = "Recency")
plot(RMFAnalysisClean$Monetary.mean, RMFAnalysisClean$Frequency, 
     ylab = "Frequency", xlab = "monetary mean")
```


```{r pressure, echo=FALSE}
par(op)

# 6 2d pairs
plot(RMFAnalysisClean, col=rainbow(7),main = "")
```

## 3D graph
```{r}
library(plotly)
fig <- plot_ly(RMFAnalysisClean, x = ~Recency, y = ~Frequency, z = ~Monetary.mean, marker = list(color = ~Monetary.mean, colorscale = c('#FFE1A1', '#683531'), showscale = TRUE))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Recency'),
                                   yaxis = list(title = 'Frequency'),
                                   zaxis = list(title = 'Monetary Value')))

fig
```

Data is spread around Recency, while it is mostly grouped around Monetary Value 0 to 1500 and Frequency 0 to 20

# Choosing the number of clusters and performing clustering

## Choosing the number of clusters
```{r}
K <- 2:15
nstart <- 5

RMFAnalysisClean_scaled <- scale(RMFAnalysisClean)
```

### WSS
```{r}
# Within Sum of Squares
set.seed(12345) 

WSS <- sapply(K, function(x) { 
  kmeans(RMFAnalysisClean_scaled, centers = x,
         nstart=nstart)$tot.withinss
  })
```

```{r}
plot(K, WSS, type="b")
```

it is an ‘elbow’, so we should perform other methods.

### ASW

```{r}
library(fpc)
# we decreased K for faster computation
K <- 2:10 
set.seed(12345)

ASW <- sapply(K, function(x) { 
  d <- dist(RMFAnalysisClean_scaled)
  cl <- kmeans(RMFAnalysisClean_scaled, centers = x,
               nstart=nstart)$cluster
  cluster.stats(d, cl)$avg.silwidth
})

plot(K, ASW, type="b")
```
maximum at K=4 and next maximum is K=5. By using the next method we will conclude K.

### Dunn Index

```{r}
set.seed(12345)

DI <- sapply(K, function(x) { 
  d <- dist(RMFAnalysisClean_scaled)
  cl <- kmeans(RMFAnalysisClean_scaled, centers = x, 
               nstart = nstart)$cluster
  cluster.stats(d, cl)$dunn
})

plot(K, DI, type="b")
```
chosen where the Dunn index is maximal and maximum is attained at K=3, next maximum is K=6.

Chosen number of clusters is 6.

# Performing clustering

### K-means clustering

```{r}
set.seed(12345)
# Manually select number of initial values and clusters
nstart <- 5
clusters <- 6

model.pc <- kmeans(RMFAnalysisClean_scaled, 
                   centers = clusters, 
                   nstart = nstart)

model.pc$centers
```

### Hierarchical clustering

```{r}
d <- dist(RMFAnalysisClean_scaled) # Euclidean distance matrix
model.hc <- hclust(d)
```

The dendrogram can be visualized and should be cut. We can specify the number of clusters that we want to achieve and cut the tree at the corresponding height:

```{r}
groups <- cutree(model.hc, k=clusters)

plot(model.hc)
rect.hclust(model.hc, k=clusters)
```
### DBSCAN

```{r}
library(dbscan)
kNNdistplot(RMFAnalysisClean_scaled, k = 6) 
abline(h = 0.7, lty = 2)
```

so it seems like a reasonable choice.

```{r}
model.dbs <- dbscan::dbscan(RMFAnalysisClean_scaled, 
                            eps = 0.7, 
                            minPts = 3)

model.dbs
```

DBSCAN identifies 8 clusters and 21 noise points. We observe the large cluster 6 and other clusters are small.

# Final result and comparison

```{r}
RMFAnalysisClean_scaled_df <- as.data.frame(scale(RMFAnalysisClean))
```

## Graphs comparisons 
### DBSCAN

```{r}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

plot(RMFAnalysisClean_scaled_df, col = cbbPalette[model.dbs$cluster + 1])
```

### Hierarchical clustering

```{r}
plot(RMFAnalysisClean_scaled_df, col=cbbPalette[groups], main = "hclust")
```

### K-means clustering

```{r}
plot(RMFAnalysisClean_scaled_df, col=cbbPalette[model.pc$cluster], main = "K-means")
```
First thing we can say is that DBSCAN performed badly. Our observations are pretty much grouped and density-based method assigned almost all of the observations to one cluster. Hierarchical and K-means clustering, so distance-based methods performed better and they divided customers in several groups based Monetary.mean and Frequency. K-means also divided in 2 groups based on Recency (green and blue groups).

## Interpreting results

To interpret the results, we will work with K-means, as it divided the data the best. We will plot the unscaled data and divide it in groups and then analyse the averages to interpret groups

### K-means clustering with unscaled data

```{r}
RMFAnalysisClean$group <- model.pc$cluster

plot(RMFAnalysisClean[, 1:3], col=cbbPalette[RMFAnalysisClean$group], 
     main = "K-means")
```

```{r}
RMFAnalysisClean_means <- data.frame(matrix(ncol = 3, nrow = 6))
colnames(RMFAnalysisClean_means) <- c("Frequency", "Monetary.mean", "Recency")

for (i in 1:6){
  RMFAnalysisClean_means$Frequency[i] <- mean(RMFAnalysisClean$Frequency[RMFAnalysisClean$group == i])
  RMFAnalysisClean_means$Monetary.mean[i] <- mean(RMFAnalysisClean$Monetary.mean[RMFAnalysisClean$group == i])
  RMFAnalysisClean_means$Recency[i] <- mean(RMFAnalysisClean$Recency[RMFAnalysisClean$group == i])
}
RMFAnalysisClean_means
```

### Interpreting groups

3rd group, which is light blue on the graph, are customers that spent the most amount of money. Monetary.mean = 4065.7404545

Averages of Frequency, Monetary.mean and Recency:
```{r}
RMFAnalysisClean_means[3,]
```
4th group, which is green on the graph, are customers that spent more than average amount of money. Monetary.mean = 941.2325562

Averages of Frequency, Monetary.mean and Recency:
```{r}
RMFAnalysisClean_means[4,]
```

2nd group, which is orange on the graph, are customers that purchase the most frequently. Frequency = 34.4871795

Averages of Frequency, Monetary.mean and Recency:
```{r}
RMFAnalysisClean_means[2,]
```

6th group, which is dark blue on the graph, are customers that purchase more frequently, than average. Frequency = 12.4109347

Averages of Frequency, Monetary.mean and Recency:
```{r}
RMFAnalysisClean_means[6,]
```

5th group, which is yellow on the graph, are customers that purchased the most recently, but do not purchase frequently. Recency = 46.003866

```{r}
#average frequency 
mean(RMFAnalysisClean$Frequency)
```

```{r}
#group 4 average frequency
RMFAnalysisClean_means$Frequency[5]
```

Averages of Frequency, Monetary.mean and Recency:
```{r}
RMFAnalysisClean_means[5,]
```
1st group, which is black on the graph, are customers that purchased the least recently. Recency = 249.0255363

Averages of Frequency, Monetary.mean and Recency:
```{r}
RMFAnalysisClean_means[1,]
```

